Clicks and Cliques

### 2015-08-26 papersoul.Rnw
add results from RF in data folder

<<data,cache=TRUE, echo=FALSE,warning=FALSE,message=FALSE>>=
# Data contain on 'sotc.csv' file were pre-processed by Xiaoque Cheng with mergeGUI. 
soul <- read.csv('data/sotc.csv', header=T)

@@

<<census,dependson='data', echo=FALSE,results='hide',warning=FALSE,message=FALSE>>=

@@

<<cluster, dependson='data',echo=FALSE,warning=FALSE, cache=TRUE>>=
# we choose 4 clusters
  set.seed(1234); km <- kmeans(soul.red2[,c(3,9:45)],centers=4)

@@

<<importance, dependson=c('data','cluster'), echo=FALSE,results='hide', cache=TRUE,warning=FALSE,message=FALSE>>=
# Fit RF on each community, to explain CCE with the constructed variables

@@

Figure \ref{map} shows the communities location, most are from the East Coast of the United States. Size of the dot represents population. Detroit and Philadelphia are the biggest communities while Aberdeen and Milledgeville are the smallest.
\begin{figure}[hbpt] 
\centering
<<mapita,dependson=c('data','census'),echo=FALSE, fig.height=6,message=FALSE,warning=FALSE,warning=FALSE>>=
lon.lat <- read.table('lon.lat.txt',header=T)
require(ggmap)
require(mapproj)

@@

The \cite{acs2011} presents a population distribution among age groups and gender. There are 23 categories for age, most of them grouping 5 years together, and in each age category the midpoint is computed as a representative age for plotting. Figure \ref{age} shows the proportion of males and females by age category across all the communities.  The general profiles are very similar for all communities. In most communities, the proportion of people between 18 and 21 is around 5\%. However, there are five communities where the proportion of people between 18 and 21 is larger, Grand Forks (13\%), Milledgeville (12\%), Tallahassee (13\%), Boulder (9\%), and State College (22\%) of the total population. Communities with a larger proportion of people over 55 years old are Palm Beach and Bradenton.
 \begin{figure}[hbpt]
<<agesex, echo=FALSE, dependson=c('census','data'),fig.height=7 >>=

@@

\begin{figure}
<<education,echo=FALSE, dependson=c('census','data')>>=
dat.aux <- melt(data=census.data,id.vars='QSB',measure.vars=c("Less.than.high.school.graduate.","High.school.graduate.","Some.college.or.associate.s.degree.","Bachelor.s.degree.or.higher.") )

@@

\begin{figure}[hbpt]
<<cluster.cen, dependson=c('data', 'census','corplot'), echo=FALSE,message=FALSE,warning=FALSE>>=

att.rf <- randomForest(Not.Attached ~ . , data=census.data[,c(2:63,66)], importance=T, ntree=2000)
aux <- dat.cor[order(dat.cor$cor), ]
aux <- subset(aux, type=='Attached')
var.cl1 <- as.character(aux$vars[c(1:10,51:61)])
d <- dist(census.data[, var.cl1])
hc <- hclust(d, method='average')
census.data$cl.att <- cutree(hc,3)

var.cl2 <- colnames(census.data)[c(3,8,12,17,21,22,28,31,48)]
d2 <- dist(census.data[, var.cl2])
hc2 <- hclust(d2, method='ward')
census.data$cl.2 <- cutree(hc2,4)

# Characterize the clusters with the numerical variables
aux <- split(census.data[,c(3,8,12,17,21,22,28,31,48)], census.data$cl.2)
aux1 <- ldply(aux, function(x) apply(x,2,mean) )
colnames(aux1) <- c('cl.2', 'Age.Fem', '1821.Yr','Bachelor', 'Unemp','Gini','Own.rate', 'Afri.pr', 'No.workers', 'New.entry' ) 
aux1$cl.id <- as.factor(aux1$cl.2)
ggparcoord(aux1[,-11], columns=-1,scale='std', groupColumn='cl.2',showPoints=T, order=c(3,4,6,10,2,5,9,7,8))  + theme(axis.text.x = element_text(angle=90, size=I(15))) +xlab('')+ylab('')+ scale_color_brewer(type='qual', palette=2, name='Group')
@
\caption{Parallel coordinate plot for community cluster means \label{cluster.cen}.  }
\end{figure}

@@

\begin{figure}[hbpt]
<<corplot, echo=FALSE, dependson=c('census','data'),cache=TRUE>>=
# With Census data we want to describe communities
levels(dat.cor2$vars) <- c('Aver.years.rent.in.community','Average.years.own.in.community','Prop.house.4.per.or.more',
                         'Prop.house.2.per.1.worker','Prop.owner.more.10.years','size.community','Unemployment.rate',
                         'Prop.unemployed','Prop.rent.more.10.years','Pro.female','Prop.house.3.per.2.worker',
                         'Prop.native.entry.prev.80','Prop.white','Income.per.capita','Prop.militar','Prop.house.3.per.no.workers',
                         'Prop.native.entry.90s','Prop.pop.bw.18.21','Prop.native.entry.80s','Prop.house.3.per.3.workers','Prop.male')

@@

qplot(data=dat.cor2, y=vars, x=cor,xlab='Correlation',ylab='', color=positive, facets=~type, size=I(3))+theme(legend.position='bottom',legend.title=element_blank(),axis.text.y = element_text( size=I(15)))
@
\caption{Correlations with attached and not attached proportion in order from top to bottom based on correlation with attached\label{corplot}. This plot shows the 10 more relevant variables with positive correlation and 10 with negative correlation with attachment.}
\end{figure}

@@

\begin{figure}[hbpt]
<<cluster.plot2, dependson='cluster',echo=FALSE, fig.height=4,message=FALSE,warning=FALSE,warning=FALSE,fig.height=6.5>>=
ind <- colnames(soul.red2) == 'cl.id'
colnames(soul.red2)[ind] <- 'Cluster.ID'

@@

\begin{figure}
<<life, dependson=c('data', 'cluster','census'), echo=FALSE,fig.height=4,message=FALSE,warning=FALSE>>=

@@

<<moving, dependson=c('data', 'cluster'), echo=FALSE,results='asis'>>=

@@

The bottom panel of Figure \ref{eco} presents the U.S. GDP growth rate in 2008, 2009, and 2010. A similar pattern is observed in both panels, suggesting the economic dimension is less related with attachment during the economic crisis. 
\begin{figure}[hbpt]
<<econ, dependson=c('data', 'cluster'), echo=FALSE,message=FALSE,warning=FALSE, fig.height=5>>=
wt <- with(subset(soul,!is.na(PROJWT)), PROJWT)
des.wt <- svydesign(ids=~1, strata=~QSB, weights=~wt, data=subset(soul,!is.na(PROJWT))) 
vars  <- svyby(~ECONOMY+CCE, ~YEAR, des.wt, svyvar, keep.var=TRUE, na.rm=T)
vars <- mutate(vars,rho = with(vars, V2/sqrt(V1*V4)),gdp = c(-0.4,-3.5,3.0))

@@

<<race, echo=FALSE, dependson=c('census', 'data' ,'cluster'), message=FALSE, fig.height=7>>= 
#relation between attached prop of white people and propo of.afro

@@

\begin{figure}[hbpt]
<<import.plot1, dependson='importance',echo=FALSE, fig.height=5,message=FALSE,warning=FALSE,warning=FALSE>>=
rankings$dim <- factor(rankings$dim)
rankings$dim <- recode(rankings$dim,"'QD2'='Residence.Years';'QD1'='Age';'Q15AB'='Lead.Int';'QD9'='Income';'Q9'='Economy';'Q8F'='Senior';'Q8D'='Children';'Q7M'='Care';'Q7L'='Leader';'Q7K'='Health';'Q7I'='Meet.People';'Q7G'='Colleges';'Q7F'='Schools';'Q7B'='Beauty';'Q6A'='5year.ahead';'Q6'='5year.ago'")

@@

The variable 5.years.ago represents the perception of the community improvement, when comparing the community today with 5 years ago. Finally, the children variable indicates if the community is a good place for families with children.
\begin{figure}[hbpt]
<< import.plot3, dependson=c('data','cluster'), echo=FALSE >>=
aux <- soul.red2
aux$QSB <- with(aux, reorder(QSB, -as.numeric(QSB) ))

@@

\begin{figure}
<<import.plot2, dependson='importance',echo=FALSE, fig.height=6, fig.width=10,message=FALSE,warning=FALSE,warning=FALSE>>=
or.fun <- function(x) -max(x)
important$dim2 <- with(important, reorder(dim, rmdg, or.fun)   )
aux <- subset(important, dim2 %in% levels(dim2)[1:10])

@@

\begin{figure}
<<import.plot4, dependson='importance',echo=FALSE, fig.height=5, fig.width=8,message=FALSE,warning=FALSE,warning=FALSE, fig.keep='all'>>=
library(ggdendro)

>>>>

<<data,echo=FALSE,warning=FALSE,message=FALSE>>=
# Procesing data sets from SOCT and Census Bureau
source('code/data_chk.R')
source('code/census_chk.R')

# Compute comunity proportion of attached people
lev <- levels(soul$QSB)
census.data %<>%
        mutate(QSB = factor(census.data$QSB, labels=lev) ) %>%
        left_join(att.pr[, c(1,2,4)], by = "QSB") %>% 
        mutate(QSB = reorder(QSB, CCEGRP2Attached))

# loading results from cluster and random forest models. 
load('data/res_cluster.Rdta')
load('data/res_rf.Rdta')
soul.red2$cl.id <- factor(km$cluster)
rankings <- ddply(subset(important, rnk < 6), .(dim, rnk), function(x) length(x$dim))
rankings$rnk <- as.factor(rankings$rnk)
levels(rankings$rnk) <- paste("R", 1:5, sep = "")
@

<<data2,cache=TRUE, echo=FALSE,warning=FALSE,message=FALSE,eval=FALSE>>=
# Data contain on 'sotc.csv' file were pre-processed by Xiaoque Cheng with mergeGUI. 
soul <- read.csv('data/sotc.csv', header=T)

@@

<<census,dependson='data', echo=FALSE,results='hide',warning=FALSE,message=FALSE,eval=FALSE>>=

@@

<<cluster, dependson='data',echo=FALSE,warning=FALSE, cache=TRUE,eval=FALSE>>=
# we choose 4 clusters
  set.seed(1234); km <- kmeans(soul.red2[,c(3,9:45)],centers=4)

@@

<<importance, dependson=c('data','cluster'), echo=FALSE,results='hide', cache=TRUE,warning=FALSE,message=FALSE,eval=FALSE>>=
# Fit RF on each community, to explain CCE with the constructed variables

@@

Figure \ref{map} shows the communities location, most are from the East Coast of the United States. Size of the dot represents population. Detroit and Philadelphia are the biggest communities while Aberdeen and Milledgeville are the smallest.
\begin{figure}[hbpt] 
\centering
<<mapita,dependson='data',echo=FALSE, fig.height=6,message=FALSE,warning=FALSE,warning=FALSE>>=
lon.lat <- read.table('lon.lat.txt',header=T)
require(ggmap)
require(mapproj)

@@

The \cite{acs2011} presents a population distribution among age groups and gender. There are 23 categories for age, most of them grouping 5 years together, and in each age category the midpoint is computed as a representative age for plotting. Figure \ref{age} shows the proportion of males and females by age category across all the communities.  The general profiles are very similar for all communities. In most communities, the proportion of people between 18 and 21 is around 5\%. However, there are five communities where the proportion of people between 18 and 21 is larger, Grand Forks (13\%), Milledgeville (12\%), Tallahassee (13\%), Boulder (9\%), and State College (22\%) of the total population. Communities with a larger proportion of people over 55 years old are Palm Beach and Bradenton.
 \begin{figure}[hbpt]
<<agesex, echo=FALSE, dependson='data',fig.height=7 >>=

@@

\begin{figure}
<<education,echo=FALSE, dependson='data'>>=
dat.aux <- melt(data=census.data,id.vars='QSB',measure.vars=c("Less.than.high.school.graduate.","High.school.graduate.","Some.college.or.associate.s.degree.","Bachelor.s.degree.or.higher.") )

@@

\begin{figure}[hbpt]
<<cluster.cen, dependson=c('data','corplot'), echo=FALSE,message=FALSE,warning=FALSE>>=

# att.rf <- randomForest(Not.Attached ~ . , data=census.data[,c(2:63,66)], importance=T, ntree=2000)
# aux <- dat.cor[order(dat.cor$cor), ]
# aux <- subset(aux, type=='Attached')
# var.cl1 <- as.character(aux$vars[c(1:10,51:61)])
# d <- dist(census.data[, var.cl1])
# hc <- hclust(d, method='average')
# census.data$cl.att <- cutree(hc,3)
# var.cl2 <- colnames(census.data)[c(3,8,12,17,21,22,28,31,48)]

var.cl2 <- colnames(census.data) %in% c("aveageF","prop1821", "Bachelor.s.degree.or.higher." , "gini" , "own.rate" , "afri.prop",'prop.new.entry','un.rate',"prop.no.workers" )
d2 <- dist(census.data[, var.cl2])
hc2 <- hclust(d2, method='ward.D')
census.data$cl.2 <- cutree(hc2,4)

# Characterize the clusters with the numerical variables
#aux <- split(census.data[,c(3,8,12,17,21,22,28,31,48)], census.data$cl.2)
aux <- split(census.data[,var.cl2], census.data$cl.2)
aux1 <- ldply(aux, function(x) apply(x,2,mean) )
colnames(aux1) <- c('cl.2', 'Age.Fem', '1821.Yr','Bachelor', 'Unemp','Gini','Own.rate', 'Afri.pr', 'No.workers', 'New.entry' )

#aux1$cl.id <- as.factor(aux1$cl.2)
ggparcoord(aux1[,-11], columns=-1,scale='std', groupColumn='cl.2',showPoints=T, order=c(3,4,6,10,2,5,9,7,8))  + theme(axis.text.x = element_text(angle=90, size=I(15))) +xlab('')+ylab('')+ scale_color_brewer(type='qual', palette=2, name='Group')
@
\caption{Parallel coordinate plot for community cluster means \label{cluster.cen}.  }
\end{figure}

@@

\begin{figure}[hbpt]
<<corplot, echo=FALSE, dependson=c('data'),cache=TRUE>>=
# With Census data we want to describe communities
cors1 <- cor(x=census.data[, 2:63], y=census.data[,64]) 
cors2 <- cor(x=census.data[,2:63], y=census.data[,65])

nm <- names(cors1[ rank(cors1) %in% c(1:10, 53:62), ])

dat.cor <- rbind(
data.frame(vars=rownames(cors1), cor=cors1[,1], positive=cors1[,1]>0, type='Attached') ,
data.frame(vars=rownames(cors2), cor=cors2[,1], positive=cors2[,1]>0, type='Not Attached') )

dat.cor2 <- subset(dat.cor, vars %in% nm)
dat.cor2$vars <- reorder(factor(dat.cor2$vars), (dat.cor2$cor) *(dat.cor2$type=='Attached'), mean)

levels(dat.cor2$vars) <-  c('Aver.years.own.in.community','size.community','Average.years.rent.in.community','Prop.owner.more.10.years','Pro.female','Prop.house.2.per.1.worker','Prop.house.4.per.or.more','Prop.rent.more.10.years','Prop.own','Unemployment.rate','Prop.1worker','Prop.bachelor','Prop.house.3.per.3.workers','Prop.native.entry.90s','Prop.native.entry.80s','Prop.militar','Prop.pop.bw.18.21','Prop.new.entry','Prop.male','Prop.2perhouse')

@@

qplot(data=dat.cor2, y=vars, x=cor,xlab='Correlation',ylab='', color=positive, facets=~type, size=I(3))+theme(legend.position='bottom',legend.title=element_blank(),axis.text.y = element_text(size=I(15)))
@
\caption{Correlations with attached and not attached proportion in order from top to bottom based on correlation with attached\label{corplot}. This plot shows the 10 more relevant variables with positive correlation and 10 with negative correlation with attachment.}
\end{figure}

@@

\begin{figure}[hbpt]
<<cluster.plot2, dependson='data',echo=FALSE, fig.height=4,message=FALSE,warning=FALSE,warning=FALSE,fig.height=6.5>>=
ind <- colnames(soul.red2) == 'cl.id'
colnames(soul.red2)[ind] <- 'Cluster.ID'

@@

\begin{figure}
<<life, dependson='data', echo=FALSE,fig.height=4,message=FALSE,warning=FALSE>>=

@@

<<moving, dependson='data', echo=FALSE,results='asis'>>=

@@

The bottom panel of Figure \ref{eco} presents the U.S. GDP growth rate in 2008, 2009, and 2010. A similar pattern is observed in both panels, suggesting the economic dimension is less related with attachment during the economic crisis. 
\begin{figure}[hbpt]
<<econ, dependson=c('data'), echo=FALSE,message=FALSE,warning=FALSE, fig.height=5>>=
wt <- with(subset(soul,!is.na(PROJWT)), PROJWT)
des.wt <- svydesign(ids=~1, strata=~QSB, weights=~wt, data=subset(soul,!is.na(PROJWT))) 
vars  <- svyby(~ECONOMY+CCE, ~YEAR, des.wt, svyvar, keep.var=TRUE, na.rm=T)
vars <- mutate(vars[, 1:5], rho = with(vars, V2/sqrt(V1*V4)),gdp = c(-0.4,-3.5,3.0))

@@

<<race, echo=FALSE, dependson=c( 'data' ), message=FALSE, fig.height=7>>= 
#relation between attached prop of white people and propo of.afro

@@

\begin{figure}[hbpt]
<<import.plot1, dependson='data',echo=FALSE, fig.height=5,message=FALSE,warning=FALSE>>=
rankings$dim <- factor(rankings$dim)
rankings$dim <- recode(rankings$dim,"'QD2'='Residence.Years';'QD1'='Age';'Q15AB'='Lead.Int';'QD9'='Income';'Q9'='Economy';'Q8F'='Senior';'Q8D'='Children';'Q7M'='Care';'Q7L'='Leader';'Q7K'='Health';'Q7I'='Meet.People';'Q7G'='Colleges';'Q7F'='Schools';'Q7B'='Beauty';'Q6A'='5year.ahead';'Q6'='5year.ago'")

@@

The variable 5.years.ago represents the perception of the community improvement, when comparing the community today with 5 years ago. Finally, the children variable indicates if the community is a good place for families with children.
\begin{figure}[hbpt]
<< import.plot3, dependson=c('data'), echo=FALSE >>=
aux <- soul.red2
aux$QSB <- with(aux, reorder(QSB, -as.numeric(QSB) ))

@@

\begin{figure}
<<import.plot2, dependson='data',echo=FALSE, fig.height=6, fig.width=10,message=FALSE,warning=FALSE,warning=FALSE>>=
or.fun <- function(x) -max(x)
important$dim2 <- with(important, reorder(dim, rmdg, or.fun)   )
aux <- subset(important, dim2 %in% levels(dim2)[1:10])

@@

\begin{figure}
<<import.plot4, dependson='data',echo=FALSE, fig.height=5, fig.width=8,message=FALSE,warning=FALSE,warning=FALSE, fig.keep='all'>>=
library(ggdendro)
